Adaptive Algorithms Employing Continuous Linear Functionals

### 2018-08-07 AutoAlgorithmLinearFunctionalDataVerC.tex
uploaded figures to Overleaf and added note.
Added optimality theorem

\section{Assuming that the Series Coefficients of the Solution Decay \\ Steadily} \label{sec:cone}

@@

\section{Optimality of the Adaptive Algorithm} \label{sec:opt}

@@

A lower bound on $\comp(\ca(\cc),\varepsilon,\rho)$ is established by   proof is accomplished by constructing fooling functions in $\cc$ with norms no greater than $\rho$.  To obtain a result that can be compared with the cost of our algorithm, we assume that 
\begin{equation} \label{lambdaRatio}
R = \sup_{k \in \naturals} \frac{\lambda_{n_{k-1}}}{\lambda_{n_k}} < \infty.
\end{equation}
This means that the $n_k$ are not too far apart with respect to the decay of $\lambda_i$ as $i \to \infty$

@@

Now we must check that $f_\pm \in \cc$. From the definition in \eqref{sumdef} it follows that for $k = 1, \ldots, j$ and $r \ge 1$
\begin{equation*}
\sigma_k(f_\pm)  \begin{cases} 
\displaystyle
\le \sigma_k(f) + \sigma_k(\eta u)\le 
c b^{k-j} + \eta\lambda_{n_{k-1}+1}\frac{b^{k-j}\norm[\cf]{u^{(k)} } }{\lambda_{n_k}}
\le b^{k-j}\left(c+\eta R \right) 
\\[1ex]
\displaystyle
\ge \sigma_k(f) - \sigma_k(\eta u)\ge 
c b^{k-j} - \eta\lambda_{n_{k-1}+1}\frac{b^{k-j}\norm[\cf]{u^{(k)} } }{\lambda_{n_k}}
\ge b^{k-j}\left(c-\eta R \right) , 
\end{cases}
\end{equation*}
Therefore, 
\begin{equation*}
\sigma_{k+r}(f_\pm)
\le b^{k+r-j}(c+\eta R) = ab^r b^{k-j}\frac{2c}{a+1}
=ab^r b^{k-j}\left(c-\eta R \right) \le a b^r \sigma_{k}(f_\pm),
\end{equation*}
which establishes that $f_\pm \in \cc$.

Although two test functions $f_\pm$ yield the same approximate solution, they have different true solutions.  In particular,
\begin{align*}
\varepsilon &\ge \max \bigl\{\norm[\cg]{S(f_+) - A^*(f_+,\varepsilon)}, \norm[\cg]{S(f_-) - A^*(f_-,\varepsilon)} \bigr\} \\
&\ge \frac 12 \bigl[\norm[\cg]{S(f^*_+) - A^*(f,\varepsilon)} + \norm[\cg]{S(f^*_-) - A^*(f,\varepsilon)}  \bigr] \\
&\qquad \qquad \text{since } A^*(f^*_\pm,\varepsilon) = A^*(f,\varepsilon) \\
&\ge \frac 12 \norm[\cg]{S(f^*_+) - S(f^*_-)} \quad \text{by the triangle inequality}\\
&\ge \frac 12 \norm[\cg]{S(f^*_+ - f^*_-)} \quad \text{since $S$ is linear}\\
&= \eta \norm[\cg]{S(u)}.
\end{align*}
Thus, we have
\begin{align*}
\varepsilon^2  & \ge \eta \norm[\cg]{S(u)}^2= 
\eta^2 \sum_{k=0}^{j} \bignorm[\cg]{S(u^{(k)})}^2  \frac{b^{2(k-j)}}{\lambda_{n_k}^2}\\
& \ge \eta^2 
\sum_{k=0}^{j} \bignorm[\cf]{u^{(k)}}^2 
b^{2(k-j)}\\
& \ge  \eta^2 b^{2(k^{*}-j)} \qquad \text{ where } k^* = \argmax_{0 \le k \le j} \bignorm[\cf]{u^{(k)}} \\
& \ge \eta^2 = \frac{(a-1)^2c^2}{(a+1)^2R^2} \\
&
=\frac{(a-1)^2 \rho^2}{(a+1)^2A^2}\left[\left(1+\frac{(a-1)^2}{(a+1)^{2} A^2 }\right)\sum_{k=1}^j \frac{b^{2(k-j)}}{\lambda_{n_{k}}^2}\right]^{-1} \\
& = \rho^2 \left[\left\{\frac{(a+1)^{2} R^2 }{(a-1)^2} + 1\right\}\sum_{k=1}^j \frac{b^{2(k-j)}}{\lambda_{n_{k}}^2}\right]^{-1} 
\end{align*}

@@

@@

\yuhannote{The figure is here in comment}
\begin{figure}
	\centering
	\includegraphics[width = 8 cm]{ProgramsImages/SimDirectSolErr.eps}
\end{figure}

>>>>

\section{Assuming a Steady Decay of the Series Coefficients of the Solution} \label{sec:cone}

@@

\section{Essential Optimality of the Adaptive Algorithm} \label{sec:opt}

@@

A lower bound on $\comp(\ca(\cc),\varepsilon,\rho)$ is established by   proof is accomplished by constructing fooling functions in $\cc$ with norms no greater than $\rho$.  To obtain a result that can be compared with the cost of our algorithm, we assume that 
\begin{equation} \label{lambdaRatio}
R = \sup_{k \in \naturals} \frac{\lambda_{n_{k-1}}}{\lambda_{n_k}} < \infty.
\end{equation}
This means that the $n_k$ are not too far apart with respect to the decay of $\lambda_i$ as $i \to \infty$.

The following theorem establishes a lower bound on the complexity of our problem for input functions in $\cc$.  After that we show that the cost of our algorithm as given in Theorem \ref{thm:compcost} is essentially no worse than this lower bound.

@@

Now we must check that $f_\pm \in \cc$. From the definition in \eqref{sumdef} it follows that for $k = 1, \ldots, j$ and $r \ge 1$,
\begin{equation*}
\sigma_k(f_\pm)  \begin{cases} 
\displaystyle
\le \sigma_k(f) + \sigma_k(\eta u)\le 
c b^{k-j} + \eta\lambda_{n_{k-1}+1}\frac{b^{k-j}\norm[\cf]{u^{(k)} } }{\lambda_{n_k}}
\le b^{k-j}\left(c+\eta R \right) 
\\[1ex]
\displaystyle
\ge \sigma_k(f) - \sigma_k(\eta u)\ge 
c b^{k-j} - \eta\lambda_{n_{k-1}+1}\frac{b^{k-j}\norm[\cf]{u^{(k)} } }{\lambda_{n_k}}
\ge b^{k-j}\left(c-\eta R \right) , 
\end{cases}
\end{equation*}
Therefore, 
\begin{equation*}
\sigma_{k+r}(f_\pm)
\le b^{k+r-j}(c+\eta R) = ab^r b^{k-j}\frac{2c}{a+1}
=ab^r b^{k-j}\left(c-\eta R \right) \le a b^r \sigma_{k}(f_\pm),
\end{equation*}
which establishes that $f_\pm \in \cc$.

Although two test functions $f_\pm$ yield the same approximate solution, they have different true solutions.  In particular,
\begin{align*}
\varepsilon &\ge \max \bigl\{\norm[\cg]{S(f_+) - A^*(f_+,\varepsilon)}, \norm[\cg]{S(f_-) - A^*(f_-,\varepsilon)} \bigr\} \\
&\ge \frac 12 \bigl[\norm[\cg]{S(f^*_+) - A^*(f,\varepsilon)} + \norm[\cg]{S(f^*_-) - A^*(f,\varepsilon)}  \bigr] \\
&\qquad \qquad \text{since } A^*(f^*_\pm,\varepsilon) = A^*(f,\varepsilon) \\
&\ge \frac 12 \norm[\cg]{S(f^*_+) - S(f^*_-)} \quad \text{by the triangle inequality}\\
&\ge \frac 12 \norm[\cg]{S(f^*_+ - f^*_-)} \quad \text{since $S$ is linear}\\
&= \eta \norm[\cg]{S(u)}.
\end{align*}
Thus, we have
\begin{align*}
\varepsilon^2  & \ge \eta \norm[\cg]{S(u)}^2= 
\eta^2 \sum_{k=0}^{j} \bignorm[\cg]{S(u^{(k)})}^2  \frac{b^{2(k-j)}}{\lambda_{n_k}^2}\\
& \ge \eta^2 
\sum_{k=0}^{j} \bignorm[\cf]{u^{(k)}}^2 
b^{2(k-j)}\\
& \ge  \eta^2 b^{2(k^{*}-j)} \qquad \text{ where } k^* = \argmax_{0 \le k \le j} \bignorm[\cf]{u^{(k)}} \\
& \ge \eta^2 = \frac{(a-1)^2c^2}{(a+1)^2R^2} \\
&
=\frac{(a-1)^2 \rho^2}{(a+1)^2A^2}\left[\left(1+\frac{(a-1)^2}{(a+1)^{2} A^2 }\right)\sum_{k=0}^j \frac{b^{2(k-j)}}{\lambda_{n_{k}}^2}\right]^{-1} \\
& \le  \rho^2 \left[\left\{\frac{(a+1)^{2} R^2 }{(a-1)^2} + 1\right\}\sum_{k=0}^j \frac{b^{2(k-j)}}{\lambda_{n_{k}}^2}\right]^{-1} 
\end{align*}

@@

The next step is to show that the cost of our algorithm is essentially no worse than that of the optimal algorithm.

\begin{theorem}
\label{thm:CostNoWorse}
Under assumption \eqref{lambdaRatio} $\cost(\tA,\rho,\varepsilon)$ is essentially no worse than $\comp(\ca(\cc),\rho,\varepsilon)$.
\end{theorem}
\begin{proof}
Let  
\begin{equation} \label{omegadef}
    \omega = \sqrt{\frac{(1 - b^2)}{a^4(1 + b^2R^2 + b^4R^4)}\left[\frac{(a+1)^{2} R^2 }{(a-1)^2} + 1\right]^{-1}}
\end{equation}
For any positive $\rho$ and $\varepsilon$, Theorem \ref{thm:compcost} says that $\cost(\tA,\rho,\varepsilon) \le n_{j^\dagger}$ where 
\begin{align*} 
j^\dagger &\le \min \left \{j \in \naturals : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^2b^2} \left[ \sum_{k=1}^{j-1} \frac{b^{2(k-j)}}{a^2\lambda_{n_{k-1}+1}^2} + \frac{1}{\lambda_{n_{j-1}+1}^2}\right]   \right\} \\
&\le \min \left \{j \in \naturals : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^4b^2} \sum_{k=1}^{j} \frac{b^{2(k-j)}}{\lambda_{n_{k-1}+1}^2} \right\} \qquad \text{since } a > 1\\
&\le \min \left \{j \in \naturals : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^4} \sum_{k=0}^{j-1} \frac{b^{2(k-j)}}{\lambda_{n_{k}+1}^2} \right\} \\
&\le \min \left \{j \in \naturals : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^4} \sum_{k=0}^{j-1} \frac{b^{2(k-j)}}{\lambda_{n_{k}}^2} \right\} \qquad \text{since } \lambda_{n_k} \ge \lambda_{n_k+1}\\
&\le \min \left \{j \in \naturals : \frac{\rho^2}{\varepsilon^2} \le \frac{(1 - b^2)}{a^4(1 + b^2R^2 + b^4R^4)} \sum_{k=0}^{j+1} \frac{b^{2(k-j)}}{\lambda_{n_{k}}^2} \right\} \qquad \text{by \eqref{lambdaRatio} } \\
&\le \min \left \{j \in \naturals : \frac{\rho^2}{\omega^2 \varepsilon^2} \le \left[\frac{(a+1)^{2} R^2 }{(a-1)^2} + 1\right] \sum_{k=0}^{j+1} \frac{b^{2(k-j)}}{\lambda_{n_{k}}^2} \right\} \qquad \text{by \eqref{omegadef} } \\
&\le \max \left \{j \in \naturals : \frac{\rho^2}{\omega^2 \varepsilon^2} > \left[\frac{(a+1)^{2} R^2 }{(a-1)^2} + 1\right] \sum_{k=0}^{j} \frac{b^{2(k-j)}}{\lambda_{n_{k}}^2} \right\} =:j^*.
\end{align*}
By Theorem \ref{thm:lowbdcomp}, $\comp(\ca(\cc),\omega\varepsilon,\rho) \ge n_{j^*}$, and by the argument above $n_{j^*} \ge n_{j^\dagger} \ge \cost(\tA,\varepsilon,\rho)$.  Thus, our algorithm is essentially no more costly than the optimal algorithm.
\end{proof}

@@

\yuhannote{The figure is here in comment} \frednote{The figure looked bad when I compiled it on my screen so I re-ran Matlab and then uploaded the new figures.}
\begin{figure}
	\centering
	\includegraphics[width = 8 cm]{ProgramsImages/SimDirectSolErr.eps}
\end{figure}
